### 2018-03-30 paper/sexSelectionSpacing-ver1.tex
Number of censored births by spell generated instead of hard-coded
For each spell a footnote now indicates how many births are observed
but not included in estimations because they occur after the chosen
cut-off point.

All spells continue until either a child is born or censoring occurs.
Censoring can happen for three reasons:
the survey takes place;
sterilization of the woman or her husband;
or too few births are observed for the method to work.
For the first spell I set censoring at 108 months (nine years) after marriage,
while for the second, third, and fourth spells censoring is at 96 months 
(eight years) after a birth can occur.
TK number of births observed after these periods.
With these cut-offs, less than one percent of observed occur after
the spell cut-off.

@@

Within the second spell, both urban and rural women show a declining
sex ratio as duration increases for the 1995--2004 period, whereas there
is little change, or possibly a small increase, with duration for the 
2005--2016 period.
These results are consistent with a situation where fertility is declining
over time leaving less room for a change in the decision to use sex selection.
Fertility decline is not immediately apparent from the second spell 
results, but the third spell shows that the likelihood of progressing from
two to third children declines by approximately 15 percentage points if
the woman has given birth to a boy.
Furthermore, as the Appendix shows, only about 25\% of urban women with 
two boys have a third child in the 2005--2016 period, 
down from close to 40\% in the 1995--2004 period. 
Furthermore, it is unlikely that a substantial number of births will take
place after the end of the spell period, which is equal to eight year, in
addition to the nine months from the previous birth to the start of the spell.
TK check these numbers with new spell lengths.

@@

TK check these numbers
The potential for changes in sex ratio within a spell means
that censoring is an important issue when trying to understand the development
in sex ratio.
This is especially the case when it comes to establishing whether 
the use of sex selection is increasing or decreasing.
Take, for example, the case of women with eight or more years of
education in the last period. 
If we look at women who started their third spell in either the
2005--2009 or the 2010--2014 periods and we treat end of 2015
as the cut-off time, then the observed sex ratio for the 2005--2009
women is 55.99\% boys while it is 55.38\% for women who started 
in the 2010--2014 period.%
\footnote{
This combines urban and rural women; the basis result 
changes little if I split by area of residence.
I chose the 2014 cut-off to ensure that both groups cover 
five years, even though NFHS-4 took place in 2015 and 2016.
}
It would, therefore, appear that the use of sex selection for
this spell is slightly decreasing over time.
This interpretation, however, ignores the role of censoring.
If we use the estimation results to predict final sex ratio
once childbearing for the third spell is over, the predicted
sex ratios are 56.10\% boys and 56.08\% boys for the 
2005--2009 and the 2010--2015 starts.
The reason for the difference is that while for the 
2005--2009 start we would have observed 7,134 birth out
of a predicted 7,451 total number of births, we would
only have observed 3,348 births out of a predicted 8,842
for the 2010--2014 group.

@@

Second, we need to understand how female labor force participation
interacts with the use of sex selection.
Increased autonomy for women, for example from better opportunities
for working outside the home, has been suggested as a way to increase
women's status and thereby lower the use of sex selection \citep{Das-Gupta2016}.
This may, however, be a double-edged sword.
On the one hand, it is an clear benefit to the women who gain  
bargaining power, and it also increases the cost of repeated 
sex-selective abortions.
On the other hand, it may further lower desired fertility, and 
everything else equal, that may lead to higher use of sex selection.
Understanding the trade-off between long-term benefits from improvements in
women's labor force participation and short-term costs from potential 
increases in sex selection is of paramount importance.

@@

@@

>>>>

All spells continue until either a child is born or censoring occurs.
Censoring can happen for three reasons:
the survey takes place;
sterilization of the woman or her husband;
or too few births are observed for the method to work.
For the first spell I set censoring because of too few births to 108 months 
(nine years) after marriage,
while for the second, third, and fourth spells censoring is at 96 months 
(eight years) after a birth can occur.
With these cut-offs, less than one percent of observed births occur after
the spell cut-off.%
\footnote{
\input{../tables/num_missed.tex}
Recall also that these numbers are totals across all 36 separate regressions.
}

@@

Within the second spell, both urban and rural women show a declining
sex ratio as duration increases for the 1995--2004 period, whereas there
is little change, or possibly a small increase, with duration for the 
2005--2016 period.
These results are consistent with a situation where fertility is declining
over time leaving less room for a change in the decision to use sex selection.
Fertility decline is not immediately apparent from the second spell 
results, but the third spell shows that the likelihood of progressing from
two to third children declines by approximately 15 percentage points if
the woman has given birth to a boy.
Furthermore, as the Appendix shows, only about 25\% of urban women with 
two boys have a third child in the 2005--2016 period, 
down from close to 60\% in the 1995--2004 period.

@@

The potential for changes in sex ratio within a spell means
that censoring is an important issue when trying to understand the 
development in sex ratio.
This is especially the case when it comes to establishing whether 
the use of sex selection is increasing or decreasing.
Take, for example, the case of women with eight or more years of
education in the last period. 
If we look at women who started their third spell in either the
2005--2009 or the 2010--2014 periods and we treat end of 2015
as the cut-off time, then the observed sex ratio for the 2005--2009
women is 55.99\% boys while it is 55.38\% for women who started 
in the 2010--2014 period.%
\footnote{
This combines urban and rural women; the basis result 
changes little if I split by area of residence.
I chose the 2014 cut-off to ensure that both groups cover 
five years, even though NFHS-4 took place in 2015 and 2016.
}
It would, therefore, appear that the use of sex selection for
this spell is slightly decreasing over time.
This interpretation, however, ignores the role of censoring.
If we use the estimation results to predict final sex ratio
once childbearing for the third spell is over, the predicted
sex ratios are 56.10\% boys and 56.08\% boys for the 
2005--2009 and the 2010--2015 starts.
The reason for the difference is that while for the 
2005--2009 start we would have observed 7,134 birth out
of a predicted 7,451 total number of births, we would
only have observed 3,348 births out of a predicted 8,842
for the 2010--2014 group.

@@

Second, we need to understand how female labor force participation
interacts with the use of sex selection.
Increased autonomy for women, for example from better opportunities
for working outside the home, has been suggested as a way to increase
women's status and thereby lower the use of sex selection \citep{Das-Gupta2016}.
This may, however, be a double-edged sword.
On the one hand, it is an clear benefit to the women who gain  
bargaining power, and it increases the cost of repeated 
sex-selective abortions because the increased duration between births
would cause a stronger disruption in labor market participation.
On the other hand, it may further lower desired fertility, and 
everything else equal, that may lead to higher use of sex selection.
Understanding the trade-off between long-term benefits from improvements in
women's labor force participation and short-term costs from potential 
increases in sex selection is of paramount importance.

@@

\section*{Appendices for Online Publication}

These appendices are intended for online publication.
They provide the descriptive statistics, additional
estimated duration tables, and graphs for all 
groups and periods.

@@

\vspace*{-0.51cm}
